---
title: "Bangkok Air Quality"
author: "Rahima Baboo"
format: html
editor: source
---

getwd()
setwd("C:/Users/Rahima/OneDrive - University of Cape Town/Documents/Rahima postgrad/GIT repos/BangkokAirQuality_repo")

system("git status")

#Calling libraries

library(tidyverse)
library(readr)
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)

#Read data
data <- read_delim("Data/bangkok-air-quality.csv", delim = ";") %>%
  select(-last_col())

head(data)
summary(data)
names(data)
typeof(data)
str(data)

#Getting things tidy

##Changing variables

names(data) <- trimws(names(data))

str(data$NO2)

data$NO2 <- as.numeric(data$ NO2)

##Removing columns and NA's

:::{.callout-note}
I am removing the columns: "PM10", "O3", and "SO2". I will only be comparing and analysing PM25 and NO2 as these are the harmful pollutants that are predominantly released from biomass burning.
:::

df <- data %>%
  select(-c(6, 7, 9)) %>%  
  drop_na("PM25", "NO2")

head(df)
print (df)
typeof(df)
str(df)

##Convert wide format data to long format

df <- read_csv(df)%>%
  pivot_longer(cols = c("PM25", "NO2"),
               names_to = "variable", values_to = "value")
               
##Date column to date type
df$Date <- as.Date(df$Date)

##Extract year and month
df$Year <- factor(year(df$Date))

df$Month <- factor(month(df$Date, label = TRUE, abbr = TRUE), levels = month.abb)
               
#Generating plots

##Boxplots for PM2.5

```{r}

library(ggplot2)

ggplot(data = df, aes(x = Month, y = PM25, fill = Month)) +
  geom_boxplot(color = "black", fill = "palegreen3") +  
  labs(x = "Month", y = "PM2.5 Concentration (µg/m³)", 
       title = "PM2.5 Concentration by Month (2021-2023)") +
  scale_fill_manual(values = rainbow(12)) +  # Change this if you want a specific color palette
  theme_bw() +
  facet_wrap(~Year, ncol = 1) +  # Separate plots for each year
  theme(plot.title = element_text(size = 17, face = "bold", hjust = 0.5),  
        axis.title.x = element_text(size = 13),  
        axis.title.y = element_text(size = 13),  
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),  # Rotate x-axis labels
        axis.text.y = element_text(size = 10)) +
  guides(fill = "none")  # Remove legend
```

##Boxplots for NO2

```{r}

library(ggplot2)

ggplot(data = df, aes(x = Month, y = NO2, fill = Month)) +
  geom_boxplot(color = "black", fill = "palegreen") +  
  labs(x = "Month", y = expression("NO"[2]~" Concentration (mol/m"^2*")"), 
       title = expression("NO"[2]~" Concentration by Month (2021-2023) in Bangkok, Thailand")) +
  scale_fill_manual(values = rainbow(12)) +  # Change this if you want a specific color palette
  theme_bw() +
  facet_wrap(~Year, ncol = 1) +  
  theme(plot.title = element_text(size = 17, face = "bold", hjust = 0.5),  
        axis.title.x = element_text(size = 13),  
        axis.title.y = element_text(size = 13),  
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),  
        axis.text.y = element_text(size = 9)) +
  guides(fill = "none")  # Remove legend
```

#Run tests

##PM2.5
###ANOVA
```{r}
#one-way ANOVA
anova.PM <- aov(df$PM25 ~ df$Month)

#ANOVA summary
summary(anova.PM)

```

###Tukey's HSD test
```{r}
tukey_test.PM <- TukeyHSD(anova.PM)

#Tukey's HSD test results
print(tukey_test.PM)

```
##NO2
###ANOVA
```{r}
#one-way ANOVA
anova.NO <- aov(df$NO2 ~ df$Month)

#ANOVA summary
summary(anova.NO)

```
###Tukey's HSD test
```{r}
tukey_test.NO <- TukeyHSD(anova.NO)

#Tukey's HSD test results
print(tukey_test.NO)

```