---
title: "Bangkok Air Quality"
author: "Rahima Baboo"
format: html
editor: source
---
Setting up my working directory and checking my Git repository status

getwd()

setwd("C:/Users/Rahima/OneDrive - University of Cape Town/Documents/Rahima postgrad/GIT repos/BangkokAirQuality_repo")

system("git status")

#Calling libraries 

**Note:** Remember to install and load the necessary packages

library(tidyverse) 
library(readr) 
library(readxl) 
library(dplyr) 
library(tidyr) 
library(ggplot2) 
library(lubridate)

#Read data 

The last column of the data file had no data so I removed it. I tidied the data into a table format using the code delim = ";" as the data were seperated by semicolons

```{r}
data <- read_delim("Data/RAW_DATA_bangkok-air-quality.csv", delim = ";") %>% select(-last_col())

head(data)

str(data)
```

#Getting things tidy

##Changing variables 

In this set of code I restructure the column titled "NO2" as there was a space in the column title and the values were not numeric.

```{r}
names(data) <- trimws(names(data))

str(data$NO2)

data$NO2 <- as.numeric(data$ NO2)
```

##Removing columns and NA's 

I am removing the columns: "PM10", "O3", and "SO2". This study will only be comparing and analysing PM25 and NO2 as these are the harmful pollutants that are predominantly released from biomass burning. I have also used drop_na to remove the gaps in the data for PM2.5 and NO2

```{r}
df <- data %>% select(-c(6, 7, 9)) %>%
 drop_na("PM25", "NO2")

head(df)

str(df)
```

##Convert wide format data to long format

```{r}
df <- read_csv(df)%>% pivot_longer(cols = c("PM25", "NO2"), names_to = "variable", values_to = "value")
```

##Date column to date type and Extract year and month

Converting the date column to date objects instead of characters and Creating new columns for year and month

```{r}

df$Date <- as.Date(df$Date)

df$Year <- factor(year(df$Date))

df$Month <- factor(month(df$Date, label = TRUE, abbr = TRUE), levels = month.abb)

```

#Generating plots

##Boxplots for PM2.5

```{r}

library(ggplot2)

boxplot.PM <- ggplot(data = df, aes(x = Month, y = PM25, fill = Month)) +
  geom_boxplot(color = "black", fill = "palegreen3") +  
  labs(x = "Month", y = "PM2.5 Concentration (µg/m³)", 
       title = "PM2.5 Concentration by Month (2021-2023)") +
  scale_fill_manual(values = rainbow(12)) +  # Change this if you want a specific color palette
  theme_bw() +
  facet_wrap(~Year, ncol = 1) +  # Separate plots for each year
  theme(plot.title = element_text(size = 17, face = "bold", hjust = 0.5),  
        axis.title.x = element_text(size = 13),  
        axis.title.y = element_text(size = 13),  
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),  # Rotate x-axis labels
        axis.text.y = element_text(size = 10)) +
  guides(fill = "none")  # Remove legend

ggsave("C:/Users/Rahima/OneDrive - University of Cape Town/Documents/Rahima postgrad/GIT repos/BangkokAirQuality_repo/Figures/PM2.5boxplot.png", 
       plot = boxplot.PM, width = 8, height = 6, dpi = 300)

print(boxplot.PM)
```

##Boxplots for NO2

```{r}

library(ggplot2)

boxplot.NO <- ggplot(data = df, aes(x = Month, y = NO2, fill = Month)) +
  geom_boxplot(color = "black", fill = "palegreen") +  
  labs(x = "Month", y = expression("NO"[2]~" Concentration (mol/m"^2*")"), 
       title = expression("NO"[2]~" Concentration by Month (2021-2023) in Bangkok, Thailand")) +
  scale_fill_manual(values = rainbow(12)) +  # Change this if you want a specific color palette
  theme_bw() +
  facet_wrap(~Year, ncol = 1) +  
  theme(plot.title = element_text(size = 17, face = "bold", hjust = 0.5),  
        axis.title.x = element_text(size = 13),  
        axis.title.y = element_text(size = 13),  
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),  
        axis.text.y = element_text(size = 9)) +
  guides(fill = "none")  # Remove legend

ggsave("C:/Users/Rahima/OneDrive - University of Cape Town/Documents/Rahima postgrad/GIT repos/BangkokAirQuality_repo/Figures/NO2boxplot.png", 
       plot = boxplot.PM, width = 8, height = 6, dpi = 300)

print(boxplot.NO)
```

#Run ANOVA tests 

I did an ANOVA test to determine if there were significant differences in the concentrations of PM2.5 and NO2 across months.

##PM2.5

```{r}
#one-way ANOVA
anova.PM <- aov(df$PM25 ~ df$Month)

#ANOVA summary
summary(anova.PM)

tidy(anova.PM)

```

##NO2

```{r}
#one-way ANOVA
anova.NO <- aov(df$NO2 ~ df$Month)

#ANOVA summary
summary(anova.NO)

tidy(anova.NO)

```
